An efficient composition of bidirectional programs by tupling and lazy updates

### 2019-11-29 paper/1.introduction.tex
add references
shrinking..
add labels to sections

Here, let $[1,2]$ be the original source \texttt{s}.
$get$ is a projection: $get$ of $phead$ picks the first element of the given original source as a view.
%\texttt{get pHead [1,2] = 1}.
After the view \texttt{v} is obtained, the user can modify the view.
In this case, the user
modified the view
from $1$ to $100$.
From this updated view, how can the source be updated?
This is ``the view update problem.''
To solve this problem, BX provides $put$, an update function on the original source.
In this example, $put$ of $phead$ constructs a new list ${[100,2]}$ from the updated view (\texttt{v'}) and the original source (\texttt{s}).
Note that the update to s' uses s and v' as input.
%\texttt{put pHead [1,2] 8 = [8,2]}

@@

First, let us see the standard evaluation method of compositions: ``not keeping any intermediate states and obtaining them by evaluation when they are needed.'' In this example, ``intermediate states'' are \texttt{i$_1$} and \texttt{i$_2$}. This method is used in a BX language, BiGUL \cite{Ko:2016:BFV:2847538.2847544,Ko:2017:ABB:3177123.3158129}. The merit of this method is clear semantics and easy to implement. The disadvantage is the number of $get$s will be quadratic when the BX programs' compositions are left associative. In the evaluation of lp3, two $get$s are required for obtaining \texttt{i$_2$} and one $get$ is required for obtaining \texttt{i$_1$}. In total, three $get$s are evaluated. We explain this method in Section 2.

\textcolor{red}{If the programs are simple like the running examples, we can transform left associative compositions to right associative. However, this is not always possible. For example, let us consider a BX 'reverse' program, $breverse$. $breverse$ is constructed by application of $bfoldr$ (the actual program is shown in Section 2) and $bsnoc$.
A part of the definition of $bfoldr$, we have $(\product{Replace}{(bfoldr \ bx)}) \circ bx$. In this program, recursion occurs on the left-hand side of composition. Because of priority, it is impossible to transform this program to right associative compositions. At the same time, we can not apply program fusion \cite{Wadler:1988:DTP:80099.80104} to this program because this BX program is dynamically produced depending on the length of the input list. Therefore, it is important to have a fast evaluation method for left associative compositions.}

For achieving this, we introduce two methods that use memoization.
The first method uses memoization simply: ``keeping all intermediate states and using them when they are needed.'' There is no quadratic $get$s in this method, therefore this method improves the runtime of many BX programs. However, a disadvantage of this is the familiar tradeoff for keeping and searching values in a table. Especially, when we use large inputs, the evaluation time will be long. We explain this method in Section 3.

@@

\vspace{2mm}
In very-well behaved \cite{} BX programs,
$put$ is a complement function for $get$.
%, $get$ can be a complement function for $put$.
\vspace{2mm}

@@

For obtaining complements, we use tupling of $put$ and $get$, and produce new function $pg$. Because $put$ produces new complements for $get$, we can shrink the size.
As an example, let us consider the previous example again: $phead \circ phead \circ phead$ in figure \ref{fig:eval-comp-phead-2}. \texttt{c$_1$} and \texttt{c$_2$} are complements. \texttt{d$_1$} and \texttt{d$_2$} are valid views for \texttt{s} and \texttt{i$_1$}. The points in this figure is two. First, after evaluation of the first $pg$, we do not need to keep the original source \texttt{s}, because all its information is in \texttt{c$_1$} and \texttt{i$_1$}. Second, complements are smaller than the intermediate states in the previous figure.
%This evaluation looks better than previous two strategies: this does not require repeated evaluation and require smaller storage than the original sources.
Actually, the simple combined $pg$ is not effective for left associative compositions, because this requires other two $put$ in the right part of the figure. To achieve efficient evaluation, we use two techniques, lazy update and lazy evaluation. We explain the second method and optimization in Section 4.

In Section 5, we show experimental results of the approaches. We discuss about related work in Section 6, and conclude in Section 7.

@@

\begin{itemize}
\item This is the first attempt for seriously considering efficiency of evaluation of BX compositions. Our methods are better than or similar to the results by the original method in BiGUL for all test cases.
\item We show that memoization is effective for achieving evaluation efficiency in BX languages.
\item In BX, as far as authors know optimization by $get$ and $put$ more tight is the first attempt.
\item Although we focus on a BX language BiGUL in this paper, these techniques can be potentially used in other BX languages.
  % \begin{itemize}
    % Although the first strategy is implemented in BiGUL \cite{}, the second and third strategies are introduced by this paper.
%  \item Thanks to introduction of the strategies, it is possible to compare the evaluation strategies.
%  \item Improvement of evaluation efficiency
%  \end{itemize}
%\item Approaches used in strategy 3
%  \begin{itemize}
%  \item
% \item Optimization tequniques by tupling, and lazy update
%  \item These tequniques can be potentially used in other BX languages
%  \end{itemize}
\end{itemize}

>>>>

Here, let $[1,2]$ be the original source \texttt{s}.
$get$ is a projection: $get$ of $phead$ picks the first element of the given original source as a view.
%\texttt{get pHead [1,2] = 1}.
After the view \texttt{v} is obtained, the user can modify the view.
In this case, the user
modified the view
from $1$ to $100$.
From this updated view, how can the source be updated?
This is ``the view update problem.''
To solve this problem, BX provides $put$, an update function on the original source.
In this example, $put$ of $phead$ constructs a new list ${[100,2]}$ from the updated view (\texttt{v'}) and the original source (\texttt{s}).
Note that the update to \texttt{s'} uses \texttt{s} and \texttt{v'} as input.
%\texttt{put pHead [1,2] 8 = [8,2]}

@@

First, let us see the standard evaluation method of compositions: ``not keeping any intermediate states and obtaining them by evaluation when they are needed.'' In this example, ``intermediate states'' are \texttt{i$_1$} and \texttt{i$_2$}. This method is used in a BX language, BiGUL \cite{Ko:2016:BFV:2847538.2847544,Ko:2017:ABB:3177123.3158129}. The merit of this method is clear semantics and easy to implement. The disadvantage is the number of $get$s will be quadratic when the BX programs' compositions are left associative. In the evaluation of lp3, two $get$s are required for obtaining \texttt{i$_2$} and one $get$ is required for obtaining \texttt{i$_1$}. In total, three $get$s are evaluated. We explain this method in Section~\ref{sect:minbigul}.

\textcolor{red}{If the BX programs are simple like the running examples, we can transform left associative compositions to right associative. However, this is not always possible. For example, let us consider a BX 'reverse' program, $breverse$. $breverse$ is constructed by application of $bfoldr$ (the actual program is shown in the last part of Section~\ref{sect:minbigul}) and $bsnoc$.
A part of the definition of $bfoldr$, we have $(\product{Replace}{(bfoldr \ bx)}) \circ bx$. In this program, recursion occurs on the left-hand side of composition. Because of priority, it is impossible to transform this program to right associative compositions. At the same time, we can not apply program fusion \cite{Wadler:1988:DTP:80099.80104} to this program because this BX program is dynamically produced depending on the length of the input list. Therefore, it is important to have a fast evaluation method for left associative compositions.}

For achieving this, we introduce two methods that use memoization.
The first method uses memoization simply: ``keeping all intermediate states and using them when they are needed.'' There is no quadratic $get$s in this method, therefore this method improves the runtime of many BX programs. However, a disadvantage of this is the familiar tradeoff for keeping and searching values in a table. Especially, when we use large inputs, the evaluation time will be long. We explain this method in Section~\cite{sect:minbigl_m}.

@@

\vspace{2mm}
In very-well behaved \cite{Foster:2007:CBT:1232420.1232424} BX,% programs,
$put$ is a complement function for $get$.
%, $get$ can be a complement function for $put$.
\vspace{2mm}

@@

For obtaining complements, we use tupling of $put$ and $get$, and produce new function $pg$. Because $put$ produces new complements for $get$, we can shrink the size.
As an example, let us consider the previous example again: $phead \circ phead \circ phead$ in figure \ref{fig:eval-comp-phead-2}. \texttt{c$_1$} and \texttt{c$_2$} are complements. \texttt{d$_1$} and \texttt{d$_2$} are valid views for \texttt{s} and \texttt{i$_1$}. The points in this figure is two. First, after evaluation of the first $pg$, we do not need to keep the original source \texttt{s}, because all its information is in \texttt{c$_1$} and \texttt{i$_1$}. Second, complements are smaller than the intermediate states in the previous figure.
%This evaluation looks better than previous two strategies: this does not require repeated evaluation and require smaller storage than the original sources.
Actually, the simple combined $pg$ is not effective for left associative compositions, because this requires other two $put$ in the right part of the figure. To achieve efficient evaluation, we use two techniques, lazy update and lazy evaluation. We explain the second method and optimization in Section~\ref{sect:xpg}.

In Section~\ref{sect:experiments}, we show experimental results of the approaches. We discuss about related work in Section~\ref{sect:related}, and conclude in Section~\ref{sect:conclusion}.

@@

\begin{itemize}
\item This is the first attempt for seriously considering efficiency of evaluation of BX compositions. Our methods are better than or similar to the results by the original method in BiGUL for all test cases.
\item We show that memoization is effective for achieving evaluation efficiency in BX languages.
\item In BX, as far as authors know optimization by $get$ and $put$ more tight is the first attempt.
\item Although we focus on a BX language BiGUL in this paper, these techniques can be potentially used in other BX languages.
  % \begin{itemize}
    % Although the first strategy is implemented in BiGUL, the second and third strategies are introduced by this paper.
%  \item Thanks to introduction of the strategies, it is possible to compare the evaluation strategies.
%  \item Improvement of evaluation efficiency
%  \end{itemize}
%\item Approaches used in strategy 3
%  \begin{itemize}
%  \item
% \item Optimization tequniques by tupling, and lazy update
%  \item These tequniques can be potentially used in other BX languages
%  \end{itemize}
\end{itemize}
